---
title: "SereneMLModeling"
author: "Serene Lee"
output: html_document
---

```{r}
library(tidyverse)
library(tidyr)
library(reshape2)
library(ggplot2)
library(dplyr)
#Unique values from partd_antibioticgeneric:
PARTD_AntibioticGeneric[!duplicated(PARTD_AntibioticGeneric$`Generic Name`), ]

#IMPORT IQVIA DATA:
iqvia1 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Iqvia Data/iqvia1data.csv")
iqvia2 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Iqvia Data/iqvia2data.csv")
iqvia3 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Iqvia Data/iqvia3data.csv")
iqvia4 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Iqvia Data/iqvia4data.csv")
iqvia5 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Iqvia Data/iqvia5data.csv")

#Filter for antibiotic drugs only:
iqvia1 <- filter(iqvia1, generic_name %in% PARTD_AntibioticGeneric$`Generic Name`)
iqvia2 <- filter(iqvia2, generic_name %in% PARTD_AntibioticGeneric$`Generic Name`)
iqvia3 <- filter(iqvia3, generic_name %in% PARTD_AntibioticGeneric$`Generic Name`)
iqvia4 <- filter(iqvia4, generic_name %in% PARTD_AntibioticGeneric$`Generic Name`)
iqvia5 <- filter(iqvia5, generic_name %in% PARTD_AntibioticGeneric$`Generic Name`)

#combine all sets into one
combinediqvia <- rbind(iqvia1, iqvia2, iqvia3, iqvia4, iqvia5)

#insert region column
combinediqvia <- combinediqvia %>%
  mutate(Region = ifelse(nppes_provider_state %in% c("IL", "IN", "MI", "OH", "WI"), "East North Central",
                         ifelse(nppes_provider_state %in% c("AL", "KY", "MS", "TN"), "East South Central",
                                ifelse(nppes_provider_state %in% c("NJ", "NY", "PA"), "Middle Atlantic",
                                       ifelse(nppes_provider_state %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY"), "Mountain",
                                              ifelse(nppes_provider_state %in% c("CT", "ME", "MA", "NH", "RI", "VT"), "New England",
                                                     ifelse(nppes_provider_state %in% c("AK", "CA", "HI", "OR", "WA"), "Pacific",
                                                            ifelse(nppes_provider_state %in% c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV"), "South Atlantic",
                                                                   ifelse(nppes_provider_state %in% c("IA", "KS", "MN", "MO", "NE", "ND", "SD"), "West North Central",
                                                                          ifelse(nppes_provider_state %in% c("AR", "LA", "OK", "TX"), "West South Central", "Region Undefined")))))))))) %>%
  select(-npi, -nppes_provider_last_org_name, -nppes_provider_first_name)

```


Normalize to pop/enrollment and age distribution:
```{r}
#Based on the enrollment data: only 15% of total medical enrollees in 2017 were under the age of 65, meaning majority are over the age of 65 - knowing this, will make a variable for aggregate population for each state under 65

#importing population data for 2017
pop_2017 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2017_populationct.csv") %>%
  select(-Year) %>%
  filter(Location != "United States") %>%
  mutate(nppes_provider_state = recode(Location, 
                           "Alabama" = "AL",
                           "Alaska" = "AK",
                           "Arizona" = "AZ",
                           "Arkansas" = "AR",
                           "California" = "CA",
                           "Colorado" = "CO",
                           "Connecticut" = "CT",
                           "Delaware" = "DE",
                           "District of Columbia" = "DC",
                           "Florida" = "FL",
                           "Georgia" = "GA",
                           "Hawaii" = "HI",
                           "Idaho" = "ID",
                           "Illinois" = "IL",
                           "Indiana" = "IN", 
                           "Iowa" = "IA",
                           "Kansas" = "KS",
                           "Kentucky" = "KY",
                           "Louisiana" = "LA", 
                           "Maine" = "ME",
                           "Maryland" = "MD",
                           "Massachusetts" = "MA",
                           "Michigan" = "MI",
                           "Minnesota" = "MN",
                           "Mississippi" = "MS",
                           "Missouri" = "MO",
                           "Montana" = "MT",
                           "Nebraska" = "NE",
                           "Nevada" = "NV",
                           "New Hampshire" = "NH",
                           "New Jersey" = "NJ",
                           "New Mexico" = "NM",
                           "New York" = "NY",
                           "North Carolina" = "NC",
                           "North Dakota" = "ND",
                           "Ohio" = "OH",
                           "Oklahoma" = "OK",
                           "Oregon" = "OR",
                           "Pennsylvania" = "PA",
                           "Rhode Island" = "RI",
                           "South Carolina" = "SC",
                           "South Dakota" = "SD",
                           "Tennessee" = "TN",
                           "Texas" = "TX",
                           "Utah" = "UT",
                           "Vermont" = "VT",
                           "Virginia" = "VA",
                           "Washington" = "WA",
                           "West Virginia" = "WV",
                           "Wisconsin" = "WI",
                           "Wyoming" = "WY",
                           "Puerto Rico" = "PR",
                           "Guam"="GU",
                           "American Samoa"="AS",
                           "Virgin Islands"="VI")) %>%
  mutate(AgeUnder65_17 = Total - X65.) %>%
  rename('Total_population_17'='Total',
         'Age65+_17'='X65.',
         'Age0-18_17'='Children.0.18',
         'Age19-25_17'='Adults.19.25',
         'Age26-34_17'='Adults.26.34',
         'Age35-54_17'='Adults.35.54',
         'Age55-64_17'='Adults.55.64') %>%
  select(-Location)

#reorder columns in pop_2017
pop_2017 <- pop_2017[c(8,7,6,9,1,2,3,4,5)]

#importing Enrollment data
Enroll_17 <- read.csv("~/Documents/School/Smith Sem 8 LAST SEM/Special Studies/Data/MedicareEnroll_Data/MedicareEnroll_2017.csv")

#Data wrangling for Enroll_17
Enroll_17 <- Enroll_17 %>%
  mutate(nppes_provider_state = recode(Location, 
                           "Alabama" = "AL",
                           "Alaska" = "AK",
                           "Arizona" = "AZ",
                           "Arkansas" = "AR",
                           "California" = "CA",
                           "Colorado" = "CO",
                           "Connecticut" = "CT",
                           "Delaware" = "DE",
                           "District of Columbia" = "DC",
                           "Florida" = "FL",
                           "Georgia" = "GA",
                           "Hawaii" = "HI",
                           "Idaho" = "ID",
                           "Illinois" = "IL",
                           "Indiana" = "IN", 
                           "Iowa" = "IA",
                           "Kansas" = "KS",
                           "Kentucky" = "KY",
                           "Louisiana" = "LA", 
                           "Maine" = "ME",
                           "Maryland" = "MD",
                           "Massachusetts" = "MA",
                           "Michigan" = "MI",
                           "Minnesota" = "MN",
                           "Mississippi" = "MS",
                           "Missouri" = "MO",
                           "Montana" = "MT",
                           "Nebraska" = "NE",
                           "Nevada" = "NV",
                           "New Hampshire" = "NH",
                           "New Jersey" = "NJ",
                           "New Mexico" = "NM",
                           "New York" = "NY",
                           "North Carolina" = "NC",
                           "North Dakota" = "ND",
                           "Ohio" = "OH",
                           "Oklahoma" = "OK",
                           "Oregon" = "OR",
                           "Pennsylvania" = "PA",
                           "Rhode Island" = "RI",
                           "South Carolina" = "SC",
                           "South Dakota" = "SD",
                           "Tennessee" = "TN",
                           "Texas" = "TX",
                           "Utah" = "UT",
                           "Vermont" = "VT",
                           "Virginia" = "VA",
                           "Washington" = "WA",
                           "West Virginia" = "WV",
                           "Wisconsin" = "WI",
                           "Wyoming" = "WY",
                           "Puerto Rico" = "PR",
                           "Guam"="GU",
                           "American Samoa"="AS",
                           "Virgin Islands"="VI")) %>%
  select(-Location)

#reorder columns in Enroll_17
Enroll_17 <- Enroll_17[c(5,1,2,3,4)]

#Merging combinediqvia/Enroll_17/pop_2017 using left outer join
combinediqvia_17 <- merge(x = combinediqvia, y = pop_2017, by = "nppes_provider_state", all.x = TRUE)
combinediqvia_17 <- merge(x = combinediqvia_17, y = Enroll_17, by = "nppes_provider_state", all.x = TRUE) %>%
  filter(Region != "Region Undefined") # for nnow taking our undefined regions

#Calculate proportion of total medicare enrollment for each state
##get rid of the commas in the data
combinediqvia_17$Total_MedicareEnroll <- as.numeric(gsub(",","",combinediqvia_17$Total_MedicareEnroll))
combinediqvia_17$Total_PartDEnroll <- as.numeric(gsub(",","",combinediqvia_17$Total_PartDEnroll))
combinediqvia_17$StandAlone_DrugPlanEnroll <- as.numeric(gsub(",","",combinediqvia_17$StandAlone_DrugPlanEnroll))
combinediqvia_17$MedicareAdv_DrugPlanEnroll <- as.numeric(gsub(",","",combinediqvia_17$MedicareAdv_DrugPlanEnroll))
##turn NAS to 0
#combinediqvia_17$Total_population_17[is.na(combinediqvia_17$Total_population_17)] <- 0
#combinediqvia_17$Total_MedicareEnroll[is.na(combinediqvia_17$Total_MedicareEnroll)] <- 0
#combinediqvia_17$Total_PartDEnroll[is.na(combinediqvia_17$Total_PartDEnroll)] <- 0
#combinediqvia_17$StandAlone_DrugPlanEnroll[is.na(combinediqvia_17$StandAlone_DrugPlanEnroll)] <- 0
#combinediqvia_17$MedicareAdv_DrugPlanEnroll[is.na(combinediqvia_17$MedicareAdv_DrugPlanEnroll)] <- 0


##calculate the percent
#combinediqvia_17 <- combinediqvia_17 %>%
#  mutate(Enroll_percapita_17 = bene_count/Total_MedicareEnroll)
#         prop_planDEnroll_17 = Total_PartDEnroll/Total_population_17,
#         prop_AgeOver65 = `Age65+_17`/Total_population_17)

##Selecting wanted variables only:
combinediqvia_17 <- combinediqvia_17 %>%
  select(-nppes_provider_city, -description_flag, -drug_name, -bene_count_ge65_suppress_flag)

```

Per capita enrollment:
```{r}
# USING TOTAL_CLAIMS
region <- combinediqvia_17 %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count),
            total_medicareEnroll = mean(Total_MedicareEnroll),
            percapita_Enroll = totalprescriptioncount/total_medicareEnroll) %>% 
  arrange(desc(percapita_Enroll))
region

ggplot(region, aes(x = reorder(Region, -percapita_Enroll), y = percapita_Enroll)) + geom_col() + labs(x = "Region", y = "Total prescription Count per capita enrollment", title = "Total Claims per capita enrollment by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

#Checking the age distribution of states
Age_over65_state <- combinediqvia_17 %>%
  mutate(percent_over65 = (`Age65+_17`/Total_population_17)*100) %>%
  group_by(nppes_provider_state) %>%
  summarise(percent_over65 = mean(percent_over65, na.rm = TRUE))

Age_over65_region <- combinediqvia_17 %>%
  mutate(percent_over65 = (`Age65+_17`/Total_population_17)*100) %>%
  group_by(Region) %>%
  summarise(percent_over65 = mean(percent_over65, na.rm = TRUE))



#boxplot of percent age 65 years older
boxplot_over65 <- combinediqvia_17 %>%
  mutate(prop_age65over_Enrolls = bene_count_ge65/bene_count,
         prop_age65over = `Age65+_17`/Total_population_17) %>%
  select(nppes_provider_state, Region, prop_age65over_Enrolls, prop_age65over)

boxplot_over65 <- na.omit(boxplot_over65) 

ggplot(boxplot_over65, aes(x=Region, y=prop_age65over_Enrolls)) + 
  geom_boxplot(outlier.colour=NA) + coord_cartesian(ylim = c(0, 1.0))

ggplot(boxplot_over65, aes(x=Region, y=prop_age65over)) + 
  geom_boxplot(outlier.colour=NA) + coord_cartesian(ylim = c(0, 1.0))

boxplot_over65 <- boxplot_over65 %>%
  group_by(Region) %>%
  summarise(avgProp_age65over = mean(prop_age65over_Enrolls))

# USING TOTAL_CLAIMS_GE65 - beneficiaries older than 65 years old 
#combinediqvia_17$total_claim_count_ge65[is.na(combinediqvia_17$total_claim_count_ge65)] <- 0
#combinediqvia_17$bene_count_ge65[is.na(combinediqvia_17$bene_count_ge65)] <- 0
#combinediqvia_17$bene_count[is.na(combinediqvia_17$bene_count)] <- 0


region <- combinediqvia_17 %>%
  mutate(percent_age65over = bene_count_ge65/bene_count) %>%
  group_by(Region) %>% 
  summarise(avg_percent = mean(percent_age65over,na.rm = TRUE),
            totalprescriptioncount = sum(total_claim_count_ge65, na.rm = TRUE),
            total_relative_medicareEnroll = avg_percent*mean(Total_MedicareEnroll, na.rm = TRUE),
            percapita_Enroll = totalprescriptioncount/total_relative_medicareEnroll) %>% 
  arrange(desc(percapita_Enroll))

ggplot(region, aes(x = reorder(Region, -percapita_Enroll), y = percapita_Enroll)) + geom_col() + labs(x = "Region", y = "Total prescription Count per capita enrollment for 65+ yrs", title = "Total Claims per capita enrollment for 65+ by Region in 2017") + theme(axis.text.x = element_text(angle = 90))
region

##Compare to original WITHOUT per capita:
region <- combinediqvia %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>%
  filter(Region != "Region Undefined")

ggplot(region, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

```

#Looking at Prescribers
```{r}
#EastSouthCentral
EastSouthCentral <- combinediqvia_17 %>%
  mutate(percent_age65over = bene_count_ge65/bene_count) %>%
  filter(Region == "East South Central") %>%
  group_by(specialty_description) %>%
  summarise(avg_percent = mean(percent_age65over,na.rm = TRUE),
            totalprescriptioncount = sum(total_claim_count_ge65, na.rm = TRUE),
            total_relative_medicareEnroll = avg_percent*mean(Total_MedicareEnroll, na.rm = TRUE),
            percapita_Enroll = totalprescriptioncount/total_relative_medicareEnroll) %>% 
  arrange(desc(percapita_Enroll))


  
```



EDA for IQVIA data:

NAs
```{r}
glimpse(combinediqvia)
summary(combinediqvia)

#about NAs
combinediqvia %>% map(~ mean(is.na(.)))

#filter for only NAs in bene_count

```

description_flag
```{r}
descrip_flag <- combinediqvia %>%
  group_by(description_flag) %>%
  summarise(count = n(),
            percent = 100*(count/1519798))

ggplot(descrip_flag, aes(description_flag, percent)) + geom_col() + coord_flip()
# majority of the specialty description is associated to Medicare specialty code/Part B, if more htan one specialty - code asociated with the largest number of services reported

```


bene_count
```{r}
#histogram displaying the distribution of beneficiaries number
distrib_bene <- ggplot(combinediqvia, aes(bene_count)) + geom_histogram(na.rm = TRUE) + scale_y_log10() + facet_wrap(~Region)
distrib_bene

#least number of beneficiaries (bene_count <= 11)
combinediqvia_leastbene <- combinediqvia %>%
  filter(bene_count <= 11) #add n/a rows next time --> insert randomized counts from 0-11?

totaliqvia2017_leastbene <- combinediqvia_leastbene %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(25)

ggplot(totaliqvia2017_leastbene, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Top Prescribed Antibiotics with the LEAST number of Medicare Beneficiaries") + theme(axis.text.x = element_text(angle = 90))

#Most number of beneficiaries (bene)
combinediqvia_mostbene <- combinediqvia %>%
  filter(bene_count > 500) #add n/a rows next time --> insert randomized counts from 0-11?

totaliqvia2017_mostbene <- combinediqvia_mostbene %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(25)

ggplot(totaliqvia2017_mostbene, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Top Prescribed Antibiotics with the MOST number of Medicare Beneficiaries") + theme(axis.text.x = element_text(angle = 90))

#In general for year 2017
totaliqvia2017 <- combinediqvia %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(25)

ggplot(totaliqvia2017, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Count of Top 25 Highest Prescribed Antibiotics in 2017") + theme(axis.text.x = element_text(angle = 90))

#In general for year 2017 (ge65)
totaliqvia2017 <- combinediqvia %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count_ge65)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(totaliqvia2017, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribed Antibiotics for 65+ in 2017") + theme(axis.text.x = element_text(angle = 90))


```

specialty_description:
```{r}
#distribution of specialty
totaliqvia2017 <- combinediqvia %>% 
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(totaliqvia2017, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribers in 2017") + theme(axis.text.x = element_text(angle = 90))

#distribution of specialty
totaliqvia2017 <- combinediqvia %>% 
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange() %>% 
  head(10)

ggplot(totaliqvia2017, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Count of Bottom 10 Prescribers in 2017") + theme(axis.text.x = element_text(angle = 90))

#distribution of specialty for only 65+
total_spec_ge65 <- combinediqvia %>% 
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count_ge65)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(total_spec_ge65, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribers for 65+ in 2017") + theme(axis.text.x = element_text(angle = 90)) #There are a lot more NA values for total_claim_count_ge65

# general specialty distribution by region
## Family Medicine
region_familymedicine <- combinediqvia %>% 
  filter(specialty_description == "Family Practice") %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_familymedicine, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Family Practice Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

## Internal Medicine
region_internalmed <- combinediqvia %>% 
  filter(specialty_description == "Internal Medicine") %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_internalmed, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Internal Medicine Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

## Nurse Practitioner
region_nursepract <- combinediqvia %>% 
  filter(specialty_description == "Nurse Practitioner") %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_nursepract, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Nurse Practitioner Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

## Dentist
region_dentist <- combinediqvia %>% 
  filter(specialty_description == "Dentist") %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_dentist, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Dentist Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

## Physician Assistant
region_pa <- combinediqvia %>% 
  filter(specialty_description == "Physician Assistant") %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_pa, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Physician Assistant Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))

#South Atlantic has most claims for the top five specialties -- followed by East North Central and Middle Atlantic, Pacific and West South Central

region <- combinediqvia %>%
  group_by(Region) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region, aes(x = reorder(Region, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Region", y = "Total Prescription Claims", title = "Count of Claims by Region in 2017") + theme(axis.text.x = element_text(angle = 90))



```

total_day_supply
```{r}
remove.packages("ggplot2") # Unisntall ggplot
install.packages("ggplot2") # Install it again
library(ggplot2)

# looking at total supply with top five prescribers
total_supply_top5 <- combinediqvia %>% 
  filter(specialty_description == "Family Practice" |
           specialty_description == "Internal Medicine" |
           specialty_description == "Nurse Practitioner" |
           specialty_description == "Dentist" |
           specialty_description == "Physician Assistant")

gglot(total_supply_top5, aes(specialty_description, total_day_supply)) + geom_boxplot()

# looking at total supply with bottom five prescribers
total_supply_bottom5 <- combinediqvia %>% 
  filter(specialty_description == "Cardiology" |
           specialty_description == "Allergy/ Immunology" |
           specialty_description == "Anesthesiology" |
           specialty_description == "Cardiac Surgery" |
           specialty_description == "Addiction Medicine")

gglot(total_supply_bottom5, aes(specialty_description, total_day_supply)) + geom_boxplot()


```


Region
```{r}
#South Atlantic:
region_southatlantic <- combinediqvia %>%
  filter(Region == "South Atlantic")
##total claims by state
region_southatlantic_state <- region_southatlantic %>%
  group_by(nppes_provider_state) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_southatlantic_state, aes(x = reorder(nppes_provider_state, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "State", y = "Total Prescription Claims", title = "Count of Toal Claims by State for South Atlantic Region in 2017") + theme(axis.text.x = element_text(angle = 90))

#East North Central:
region_eastnorthcentral <- combinediqvia %>%
  filter(Region == "East North Central")
##total claims by state
region_eastnorthcentral_state <- region_eastnorthcentral %>%
  group_by(nppes_provider_state) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount))

ggplot(region_eastnorthcentral_state, aes(x = reorder(nppes_provider_state, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "State", y = "Total Prescription Claims", title = "Count of Toal Claims by State for East North Central Region in 2017") + theme(axis.text.x = element_text(angle = 90))

```

State
Florida:
```{r}
#Florida:
state_FL <- combinediqvia %>%
  filter(nppes_provider_state == "FL")
#summary(state_FL)

#Total claims in general for generic_name
state_FL_total <- state_FL %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(state_FL_total, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribed Antibiotics in Florida 2017") + theme(axis.text.x = element_text(angle = 90))

#Total Claims for year 2017 (ge65)
state_FL_ge65 <- state_FL %>% 
  group_by(generic_name) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count_ge65)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(state_FL_ge65, aes(x = reorder(generic_name, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Generic Antibiotic Name", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribed Antibiotics in Florida for 65+ in 2017") + theme(axis.text.x = element_text(angle = 90))

#What is the specialty distribution

state_FL_spec <- state_FL %>% 
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count)) %>% 
  arrange(desc(totalprescriptioncount)) %>% 
  head(10)

ggplot(state_FL_spec, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Count of Top 10 Highest Prescribers in 2017") + theme(axis.text.x = element_text(angle = 90))

state_FL_spec <- state_FL %>% 
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count_ge65)) %>% 
  arrange(desc(totalprescriptioncount)) %>%
  head(5)

ggplot(state_FL_spec, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Count of Top  Highest Prescribers for 65+ in 2017") + theme(axis.text.x = element_text(angle = 90))

#Dig deeper for ge65 and tigecycline
state_FL_tygecycline <- state_FL %>%
  filter(generic_name == "TIGECYCLINE") %>%
  group_by(specialty_description) %>% 
  summarise(totalprescriptioncount = sum(total_claim_count_ge65)) %>% 
  arrange(desc(totalprescriptioncount)) %>%
  head(10)

ggplot(state_FL_spec, aes(x = reorder(specialty_description, -totalprescriptioncount), y = totalprescriptioncount)) + geom_col() + labs(x = "Prescriber Specialty", y = "Total Prescription Claims", title = "Top Prescribers for Tigecycline for 65+ in Florida in 2017") + theme(axis.text.x = element_text(angle = 90))

```



```{r}

resistance_state <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_States.csv")

# 2) 
resistance_region <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_Regional.csv")

# 3)
resistance_national <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_National.csv")

```


#Medicaid:
##Quinolones/Fluoroquinolones antibiotic drugs:
```{r}
#Generic name of drug (Referenced GoodRx): Ciprofloxacin HCl, Ciprofloxacin, Ciprofloxacin/Hydrocortisone,
#Ciprofloxacin In 5 % Dextrose, Ciprofloxacin HCl/Dexameth, Ciprofloxacin Lactate, Ciprofloxacin/Ciprofloxa HCl, Ciprofloxacin HCl/Fluocinolone, Levofloxacin, Levofloxacin In Dextrose 5 %, Ofloxacin, Moxifloxacin HCl, Moxifloxacin-Sod.Chloride(Iso), Moxifloxacin/Sod.Ace,Sul/Water, Besifloxacin HCl, Delafloxacin Meglumine, Gatifloxacin

# Create a new data set with only drugs under the quinolones antibiotic class:
Medicaid_Quinolone <- Medicaid_SpendingUtilization %>%
  filter(GenericName == "Ciprofloxacin HCl"|
           GenericName == "Ciprofloxacin"|
           GenericName == "Ciprofloxacin/Hydrocortisone"|
           GenericName == "Ciprofloxacin In 5 % Dextrose"|
           GenericName == "Ciprofloxacin HCl/Dexameth"|
           GenericName == "Ciprofloxacin Lactate"|
           GenericName == "Ciprofloxacin/Ciprofloxa HCl"|
           GenericName == "Ciprofloxacin HCl/Fluocinolone"|
           GenericName == "Levofloxacin"|
           GenericName == "Levofloxacin In Dextrose 5 %"|
           GenericName == "Ofloxacin"|
           GenericName == "Moxifloxacin HCl"|
           GenericName == "Moxifloxacin-Sod.Chloride(Iso)"|
           GenericName == "Moxifloxacin/Sod.Ace,Sul/Water"|
           GenericName == "Besifloxacin HCl"|
           GenericName == "Delafloxacin Meglumine"|
           GenericName == "Gatifloxacin")
View(Medicaid_Quinolone)

# EDA to look at trends for spending:
ggplot(data = Medicaid_Quinolone, aes(x = Year, y = TotalSpending)) + geom_point()
ggplot(data = Medicaid_Quinolone, aes(x = Year, y = TotalClaims, col = GenericName)) + geom_jitter()
ggplot(data = Medicaid_Quinolone, aes(x = log10(TotalClaims))) + geom_histogram() + facet_wrap(~Year)
```

#Regional Resistance Data:
```{r}
# Refreshing Regional resistance data
Resist_Regional <- resistance_region %>%
  select(-Lower95CI, -Upper95CI, -PercentResistant)

#Resistance Trends by Region
Resist_Regional <- Resist_Regional %>%
  filter(EventYear != "All Years")

#Change data type
Resist_Regional$NumberTested <- as.numeric(as.character(Resist_Regional$NumberTested))
Resist_Regional$NumberResistant <- as.numeric(as.character(Resist_Regional$NumberResistant))
#Resist_Regional$PercentResistant <- as.numeric(as.character(Resist_Regional$PercentResistant))
Resist_Regional$AgeCategory <- as.character(Resist_Regional$AgeCategory)

# Inserting the average of NumberResistant and NumberTested:
Resist_Regional <- Resist_Regional %>%
  filter(NumberResistant != "Insufficient Data" & NumberResistant != "None Tested" & EventYear != "All Years")

# Plot to see trends in region:
#ggplot(data = Resist_Regional, aes(x = EventYear, y = PercentResistant, col = AgeCategory)) + geom_col()
#ggplot(data = Resist_Regional, aes(x = EventYear, y = PercentResistant)) + geom_col() + facet_wrap(~Region)

```

# Calculate Percent Resistant
```{r}
# Tidy format for Resist_Regional data and take out ALL HAIS and ALL AGES
Resist_Regional <- Resist_Regional %>%
  gather(MetricType, MetricNumber, NumberTested:NumberResistant) %>%
  filter(EventType != "All HAIs", AgeCategory != "All Ages")

# Plot NumberTested and NumberResistant Data
ggplot(data = Resist_Regional, 
       aes(x = EventYear, y = MetricNumber, fill = MetricType)) + 
  geom_col(position = position_dodge(width = 0.5)) + 
  facet_wrap(~Region)

# Regroup data for Calculated PercentResistant Data
Resist_Regional_CalculateResPercent <- Resist_Regional %>%
  group_by(Region, EventYear, MetricType) %>%
  summarize(TotalMetricNumber = sum(MetricNumber))

# Revert Dataset to Wide form
Wide_data_ResistPercent <- spread(Resist_Regional_CalculateResPercent,
                                  MetricType,
                                  TotalMetricNumber)
Wide_data_ResistPercent <- Wide_data_ResistPercent %>%
  mutate(Percent_Resistant = (NumberResistant/NumberTested)*100)

#Wide_data_ResistPercent <- melt(Resist_Regional_CalculateResPercent,
#                                id.vars = c("Region", "EventYear"),
#                                variable.name = "MetricType",
#                                value.name = "TotalMetricNumber")

#Wide_data_ResistPercent <- reshape(Resist_Regional_CalculateResPercent,
#                                   idvar = c("Region", "EventYear"),
#                                   timevar = "MetricType",
#                                   direction = "wide")

# Plot Percent Resistant Data
ggplot(data = Wide_data_ResistPercent, 
       aes(x = EventYear, y = Percent_Resistant)) + 
  geom_col() + 
  facet_wrap(~Region)

# Facet with Years
  
```

#Trying other variables: Age Category
```{r}
#Regroup Resist_Regional by AgeCategory:
Resist_Regional <- resistance_region %>%
  select(-Lower95CI, -Upper95CI, -PercentResistant)

#Resistance Trends by Region
Resist_Regional <- Resist_Regional %>%
  filter(NumberResistant != "Insufficient Data" & NumberResistant != "None Tested" & EventYear != "All Years")

#Change data type to numerical
Resist_Regional$NumberTested <- as.numeric(as.character(Resist_Regional$NumberTested))
Resist_Regional$NumberResistant <- as.numeric(as.character(Resist_Regional$NumberResistant))

# Tidy format for Resist_Regional data 
Resist_Regional <- Resist_Regional %>%
  gather(MetricType, MetricNumber, NumberTested:NumberResistant) %>%
  filter(EventType != "All HAIs", AgeCategory != "All Ages")

# Regroup with age and year
Resist_Regional_byAge <- Resist_Regional %>%
  group_by(Region, EventYear, MetricType, AgeCategory) %>%
  summarize(TotalMetricNumber = sum(MetricNumber))

# Revert Dataset to Wide form
Wide_AgeCategory <- spread(Resist_Regional_byAge,
                                  MetricType,
                                  TotalMetricNumber)
Wide_AgeCategory <- Wide_AgeCategory %>%
  mutate(Percent_Resistant = (NumberResistant/NumberTested)*100)

# Plot Percent Resistance by AgeCaetgory (bar):
ggplot(data = Wide_AgeCategory, 
       aes(x = EventYear, y = Percent_Resistant, fill = AgeCategory)) + 
  geom_col(position = "dodge") + 
  facet_wrap(~Region)
# Plot NumberTested by AgeCaetgory (bar):
ggplot(data = Wide_AgeCategory, 
       aes(x = EventYear, y = NumberTested, fill = AgeCategory)) + 
  geom_col(position = "dodge") +
  facet_wrap(~Region)

# Plot Percent Resistance by AgeCaetgory (line):
Resist_Regional_byAge <- Resist_Regional %>%
  group_by(EventYear, MetricType, AgeCategory) %>%
  summarize(TotalMetricNumber = sum(MetricNumber))

Wide_AgeCategory <- spread(Resist_Regional_byAge,
                                  MetricType,
                                  TotalMetricNumber)
Wide_AgeCategory <- Wide_AgeCategory %>%
  mutate(Percent_Resistant = (NumberResistant/NumberTested)*100)

Wide_AgeCategory <- Wide_AgeCategory %>%
  filter(Percent_Resistant != 0)

#ggplot(data = Wide_AgeCategory, 
#       aes(x = EventYear, y = Percent_Resistant, col = AgeCategory)) + 
#  geom_line() #Don't know why it's not plotting



```

#Trying out some statistical tests
```{r}
# Chi Square:
# categorical variables: Phenotype, Region, EventType, EventYear, EventCategory


#Regroup Resist_Regional by AgeCategory:
Resist_Regional <- resistance_region %>%
  select(-Lower95CI, -Upper95CI, -PercentResistant)

#Resistance Trends by Region
Resist_Regional <- Resist_Regional %>%
  filter(NumberResistant != "Insufficient Data" & NumberResistant != "None Tested" & EventYear != "All Years")

#Change data type to numerical
Resist_Regional$NumberTested <- as.numeric(as.character(Resist_Regional$NumberTested))
Resist_Regional$NumberResistant <- as.numeric(as.character(Resist_Regional$NumberResistant))

# Tidy format for Resist_Regional data 
Resist_Regional <- Resist_Regional %>%
  gather(MetricType, MetricNumber, NumberTested:NumberResistant) %>%
  filter(EventType != "All HAIs", AgeCategory != "All Ages")


#Rearrange
Resist_Regional_test <- Resist_Regional %>%
  group_by(Phenotype, Region, MetricType, AgeCategory) %>%
  summarize(TotalMetricNumber = sum(MetricNumber))

# Revert Dataset to Wide form
Wide_test <- spread(Resist_Regional_test,
                                  MetricType,
                                  TotalMetricNumber)
Wide_test <- Wide_test %>%
  mutate(Percent_Resistant = (NumberResistant/NumberTested)*100)


```

Merging Population and Age distrubution data:
```{r}
# importing age distrib data:
agedistrib_2011 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2011_agedistribution.csv")
agedistrib_2012 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2012_agedistribution.csv")
agedistrib_2013 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2013_agedistribution.csv")
agedistrib_2014 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2014_agedistribution.csv")
# importing population data:
pop_2011 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2011_populationct.csv") %>%
  select(-Total)
pop_2012 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2012_populationct.csv") %>%
  select(-Total)
pop_2013 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2013_populationct.csv") %>%
  select(-Total)
pop_2014 <- read.csv("~/Documents/GitHub/DoritLab_AntibioticTrends/Data/2014_populationct.csv") %>%
  select(-Total)

#Splitting up the regional
resistance_2011 <- resistance_state %>%
  filter(EventYear == 2011,
         AgeCategory != "All Ages") %>%
  select(-Lower95CI, -Upper95CI) %>%
  mutate(Location = recode(State, 
                        "AL" = "Alabama",
                        "AK" = "Alaska",
                        "AZ" = "Arizona",
                        "AR" = "Arkansas",
                        "CA" = "California",
                        "CO" = "Colorado",
                        "CT" = "Connecticut",
                        "DE" = "Delaware",
                        "DC" = "District Of Columbia",
                        "FL" = "Florida",
                        "GA" = "Georgia",
                        "HI" = "Hawaii",
                        "ID" = "Idaho",
                        "IL" = "Illinois",
                        "IN" = "Indiana",
                        "IA" = "Iowa",
                        "KS" = "Kansas",
                        "KY" = "Kentucky",
                        "LA" = "Louisiana",
                        "ME" = "Maine",
                        "MD" = "Maryland",
                        "MA" = "Massachusetts",
                        "MI" = "Michigan",
                        "MN" = "Minnesota",
                        "MS" = "Mississippi",
                        "MO" = "Missouri",
                        "MT" = "Montana",
                        "NE" = "Nebraska",
                        "NV" = "Nevada",
                        "NH" = "New Hampshire",
                        "NJ" = "New Jersey",
                        "NM" = "New Mexico",
                        "NY" = "New York",
                        "NC" = "North Carolina",
                        "ND" = "North Dakota",
                        "OH" = "Ohio",
                        "OK" = "Oklahoma",
                        "OR" = "Oregon",
                        "PA" = "Pennsylvania",
                        "RI" = "Rhode Island",
                        "SC" = "South Carolina",
                        "SD" = "South Dakota",
                        "TN" = "Tennessee",
                        "TX" = "Texas",
                        "UT" = "Utah",
                        "VT" = "Vermont",
                        "VA" = "Virginia",
                        "WA" = "Washington",
                        "WV" = "West Virginia",
                        "WI" = "Wisconsin",
                        "WY" = "Wyoming")) %>%
  select(-State, - PercentResistant)

# Rearrange pop_2011 and agedistrib_2011 to be tidy
pop_2011 <- pop_2011 %>%
  gather(AgeCategory, Population, Children.0.18:X65.)
agedistrib_2011 <- agedistrib_2011 %>%
  gather(AgeCategory, AgeDistribution, Children.0.18:X65.)

# Combining pop_2011 and agedistrib_2011
pop_2011 <- pop_2011 %>%
  left_join(agedistrib_2011, by = c("Location", "Year", "AgeCategory"))

# Regrouping AgeCategory to :
pop_2011 <- pop_2011 %>%
  mutate(AgeCategory1 = recode(AgeCategory, "Children.0.18" = "0-18",
                              "Adults.19.25" = "19-64",
                              "Adults.26.34" = "19-64",
                              "Adults.35.54" = "19-64",
                              "Adults.55.64" = "19-64",
                              "X65." = "65+")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1,
         EventYear = Year) %>%
  group_by(Location, EventYear, AgeCategory) %>%
  summarize(pop_count = sum(Population),
            agedistribution = sum(AgeDistribution))

#pop_2011[c("Location", "EventYear", "AgeCategory","Population", "AgeDistribution")] <- reordering failed

# rename resistance_2011 agecategory values
resistance_2011$NumberTested <- as.numeric(as.character(resistance_2011$NumberTested))
resistance_2011$NumberResistant <- as.numeric(as.character(resistance_2011$NumberResistant))

resistance_2011 <- resistance_2011 %>%
  mutate(AgeCategory1 = recode(AgeCategory, 
                              "<1" = "0-18",
                              " 1-18" = "0-18")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1) %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, NumberTested, NumberResistant) %>%
  summarize(numberTested = sum(NumberTested),
            numberResistant = sum(NumberResistant))

resistance_2011 <- resistance_2011 %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, numberTested, numberResistant) %>%
  select(-NumberResistant, -NumberTested)


# merging pop_2011 to resistance_2011:
resistance_2011$EventYear <- as.character(resistance_2011$EventYear)
pop_2011$EventYear <- as.character(pop_2011$EventYear)

resistance_2011 <- resistance_2011 %>%
  left_join(pop_2011, by = c("Location", "EventYear", "AgeCategory"))

```

```{r}
## YEAR 2012

resistance_2012 <- resistance_state %>%
  filter(EventYear == 2012,
         AgeCategory != "All Ages") %>%
  select(-Lower95CI, -Upper95CI) %>%
  mutate(Location = recode(State, 
                        "AL" = "Alabama",
                        "AK" = "Alaska",
                        "AZ" = "Arizona",
                        "AR" = "Arkansas",
                        "CA" = "California",
                        "CO" = "Colorado",
                        "CT" = "Connecticut",
                        "DE" = "Delaware",
                        "DC" = "District Of Columbia",
                        "FL" = "Florida",
                        "GA" = "Georgia",
                        "HI" = "Hawaii",
                        "ID" = "Idaho",
                        "IL" = "Illinois",
                        "IN" = "Indiana",
                        "IA" = "Iowa",
                        "KS" = "Kansas",
                        "KY" = "Kentucky",
                        "LA" = "Louisiana",
                        "ME" = "Maine",
                        "MD" = "Maryland",
                        "MA" = "Massachusetts",
                        "MI" = "Michigan",
                        "MN" = "Minnesota",
                        "MS" = "Mississippi",
                        "MO" = "Missouri",
                        "MT" = "Montana",
                        "NE" = "Nebraska",
                        "NV" = "Nevada",
                        "NH" = "New Hampshire",
                        "NJ" = "New Jersey",
                        "NM" = "New Mexico",
                        "NY" = "New York",
                        "NC" = "North Carolina",
                        "ND" = "North Dakota",
                        "OH" = "Ohio",
                        "OK" = "Oklahoma",
                        "OR" = "Oregon",
                        "PA" = "Pennsylvania",
                        "RI" = "Rhode Island",
                        "SC" = "South Carolina",
                        "SD" = "South Dakota",
                        "TN" = "Tennessee",
                        "TX" = "Texas",
                        "UT" = "Utah",
                        "VT" = "Vermont",
                        "VA" = "Virginia",
                        "WA" = "Washington",
                        "WV" = "West Virginia",
                        "WI" = "Wisconsin",
                        "WY" = "Wyoming")) %>%
  select(-State, - PercentResistant)

pop_2012 <- pop_2012 %>%
  gather(AgeCategory, Population, Children.0.18:X65.)
agedistrib_2012 <- agedistrib_2012 %>%
  gather(AgeCategory, AgeDistribution, Children.0.18:X65.)

pop_2012 <- pop_2012 %>%
  left_join(agedistrib_2012, by = c("Location", "Year", "AgeCategory"))

pop_2012 <- pop_2012 %>%
  mutate(AgeCategory1 = recode(AgeCategory, "Children.0.18" = "0-18",
                              "Adults.19.25" = "19-64",
                              "Adults.26.34" = "19-64",
                              "Adults.35.54" = "19-64",
                              "Adults.55.64" = "19-64",
                              "X65." = "65+")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1,
         EventYear = Year) %>%
  group_by(Location, EventYear, AgeCategory) %>%
  summarize(pop_count = sum(Population),
            agedistribution = sum(AgeDistribution))

resistance_2012$NumberTested <- as.numeric(as.character(resistance_2012$NumberTested))
resistance_2012$NumberResistant <- as.numeric(as.character(resistance_2012$NumberResistant))

resistance_2012 <- resistance_2012 %>%
  mutate(AgeCategory1 = recode(AgeCategory, 
                              "<1" = "0-18",
                              " 1-18" = "0-18")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1) %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, NumberTested, NumberResistant) %>%
  summarize(numberTested = sum(NumberTested),
            numberResistant = sum(NumberResistant))

resistance_2012 <- resistance_2012 %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, numberTested, numberResistant) %>%
  select(-NumberResistant, -NumberTested)


resistance_2012$EventYear <- as.character(resistance_2012$EventYear)
pop_2012$EventYear <- as.character(pop_2012$EventYear)

resistance_2012 <- resistance_2012 %>%
  left_join(pop_2012, by = c("Location", "EventYear", "AgeCategory"))

## YEAR 2013


resistance_2013 <- resistance_state %>%
  filter(EventYear == 2013,
         AgeCategory != "All Ages") %>%
  select(-Lower95CI, -Upper95CI) %>%
  mutate(Location = recode(State, 
                        "AL" = "Alabama",
                        "AK" = "Alaska",
                        "AZ" = "Arizona",
                        "AR" = "Arkansas",
                        "CA" = "California",
                        "CO" = "Colorado",
                        "CT" = "Connecticut",
                        "DE" = "Delaware",
                        "DC" = "District Of Columbia",
                        "FL" = "Florida",
                        "GA" = "Georgia",
                        "HI" = "Hawaii",
                        "ID" = "Idaho",
                        "IL" = "Illinois",
                        "IN" = "Indiana",
                        "IA" = "Iowa",
                        "KS" = "Kansas",
                        "KY" = "Kentucky",
                        "LA" = "Louisiana",
                        "ME" = "Maine",
                        "MD" = "Maryland",
                        "MA" = "Massachusetts",
                        "MI" = "Michigan",
                        "MN" = "Minnesota",
                        "MS" = "Mississippi",
                        "MO" = "Missouri",
                        "MT" = "Montana",
                        "NE" = "Nebraska",
                        "NV" = "Nevada",
                        "NH" = "New Hampshire",
                        "NJ" = "New Jersey",
                        "NM" = "New Mexico",
                        "NY" = "New York",
                        "NC" = "North Carolina",
                        "ND" = "North Dakota",
                        "OH" = "Ohio",
                        "OK" = "Oklahoma",
                        "OR" = "Oregon",
                        "PA" = "Pennsylvania",
                        "RI" = "Rhode Island",
                        "SC" = "South Carolina",
                        "SD" = "South Dakota",
                        "TN" = "Tennessee",
                        "TX" = "Texas",
                        "UT" = "Utah",
                        "VT" = "Vermont",
                        "VA" = "Virginia",
                        "WA" = "Washington",
                        "WV" = "West Virginia",
                        "WI" = "Wisconsin",
                        "WY" = "Wyoming")) %>%
  select(-State, - PercentResistant)

pop_2013 <- pop_2013 %>%
  gather(AgeCategory, Population, Children.0.18:X65.)
agedistrib_2013 <- agedistrib_2013 %>%
  gather(AgeCategory, AgeDistribution, Children.0.18:X65.)

pop_2013 <- pop_2013 %>%
  left_join(agedistrib_2013, by = c("Location", "Year", "AgeCategory"))

pop_2013 <- pop_2013 %>%
  mutate(AgeCategory1 = recode(AgeCategory, "Children.0.18" = "0-18",
                              "Adults.19.25" = "19-64",
                              "Adults.26.34" = "19-64",
                              "Adults.35.54" = "19-64",
                              "Adults.55.64" = "19-64",
                              "X65." = "65+")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1,
         EventYear = Year) %>%
  group_by(Location, EventYear, AgeCategory) %>%
  summarize(pop_count = sum(Population),
            agedistribution = sum(AgeDistribution))

resistance_2013$NumberTested <- as.numeric(as.character(resistance_2013$NumberTested))
resistance_2013$NumberResistant <- as.numeric(as.character(resistance_2013$NumberResistant))

resistance_2013 <- resistance_2013 %>%
  mutate(AgeCategory1 = recode(AgeCategory, 
                              "<1" = "0-18",
                              " 1-18" = "0-18")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1) %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, NumberTested, NumberResistant) %>%
  summarize(numberTested = sum(NumberTested),
            numberResistant = sum(NumberResistant))

resistance_2013 <- resistance_2013 %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, numberTested, numberResistant) %>%
  select(-NumberResistant, -NumberTested)


resistance_2013$EventYear <- as.character(resistance_2013$EventYear)
pop_2013$EventYear <- as.character(pop_2013$EventYear)

resistance_2013 <- resistance_2013 %>%
  left_join(pop_2013, by = c("Location", "EventYear", "AgeCategory"))

## YEAR 2014


resistance_2014 <- resistance_state %>%
  filter(EventYear == 2014,
         AgeCategory != "All Ages") %>%
  select(-Lower95CI, -Upper95CI) %>%
  mutate(Location = recode(State, 
                        "AL" = "Alabama",
                        "AK" = "Alaska",
                        "AZ" = "Arizona",
                        "AR" = "Arkansas",
                        "CA" = "California",
                        "CO" = "Colorado",
                        "CT" = "Connecticut",
                        "DE" = "Delaware",
                        "DC" = "District Of Columbia",
                        "FL" = "Florida",
                        "GA" = "Georgia",
                        "HI" = "Hawaii",
                        "ID" = "Idaho",
                        "IL" = "Illinois",
                        "IN" = "Indiana",
                        "IA" = "Iowa",
                        "KS" = "Kansas",
                        "KY" = "Kentucky",
                        "LA" = "Louisiana",
                        "ME" = "Maine",
                        "MD" = "Maryland",
                        "MA" = "Massachusetts",
                        "MI" = "Michigan",
                        "MN" = "Minnesota",
                        "MS" = "Mississippi",
                        "MO" = "Missouri",
                        "MT" = "Montana",
                        "NE" = "Nebraska",
                        "NV" = "Nevada",
                        "NH" = "New Hampshire",
                        "NJ" = "New Jersey",
                        "NM" = "New Mexico",
                        "NY" = "New York",
                        "NC" = "North Carolina",
                        "ND" = "North Dakota",
                        "OH" = "Ohio",
                        "OK" = "Oklahoma",
                        "OR" = "Oregon",
                        "PA" = "Pennsylvania",
                        "RI" = "Rhode Island",
                        "SC" = "South Carolina",
                        "SD" = "South Dakota",
                        "TN" = "Tennessee",
                        "TX" = "Texas",
                        "UT" = "Utah",
                        "VT" = "Vermont",
                        "VA" = "Virginia",
                        "WA" = "Washington",
                        "WV" = "West Virginia",
                        "WI" = "Wisconsin",
                        "WY" = "Wyoming")) %>%
  select(-State, - PercentResistant)

pop_2014 <- pop_2014 %>%
  gather(AgeCategory, Population, Children.0.18:X65.)
agedistrib_2014 <- agedistrib_2014 %>%
  gather(AgeCategory, AgeDistribution, Children.0.18:X65.)

pop_2014 <- pop_2014 %>%
  left_join(agedistrib_2014, by = c("Location", "Year", "AgeCategory"))

pop_2014 <- pop_2014 %>%
  mutate(AgeCategory1 = recode(AgeCategory, "Children.0.18" = "0-18",
                              "Adults.19.25" = "19-64",
                              "Adults.26.34" = "19-64",
                              "Adults.35.54" = "19-64",
                              "Adults.55.64" = "19-64",
                              "X65." = "65+")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1,
         EventYear = Year) %>%
  group_by(Location, EventYear, AgeCategory) %>%
  summarize(pop_count = sum(Population),
            agedistribution = sum(AgeDistribution))

resistance_2014$NumberTested <- as.numeric(as.character(resistance_2014$NumberTested))
resistance_2014$NumberResistant <- as.numeric(as.character(resistance_2014$NumberResistant))

resistance_2014 <- resistance_2014 %>%
  mutate(AgeCategory1 = recode(AgeCategory, 
                              "<1" = "0-18",
                              " 1-18" = "0-18")) %>%
  select(-AgeCategory) %>%
  rename(AgeCategory = AgeCategory1) %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, NumberTested, NumberResistant) %>%
  summarize(numberTested = sum(NumberTested),
            numberResistant = sum(NumberResistant))

resistance_2014 <- resistance_2014 %>%
  group_by(Phenotype, EventType, Location, EventYear, 
           AgeCategory, numberTested, numberResistant) %>%
  select(-NumberResistant, -NumberTested)


resistance_2014$EventYear <- as.character(resistance_2014$EventYear)
pop_2014$EventYear <- as.character(pop_2014$EventYear)

resistance_2014 <- resistance_2014 %>%
  left_join(pop_2014, by = c("Location", "EventYear", "AgeCategory"))

```

MERGING 2011-2014 Data
```{r}
#merging rows

resistance_total <- rbind(resistance_2011, resistance_2012,
                          resistance_2013, resistance_2014)

```


Multiple Imputation:
```{r}
# removing all HAIs:
resistance_total <- resistance_total %>%
  filter(EventType != "All HAIs")

resistance_total$Phenotype <- as.character(resistance_total$Phenotype)
resistance_total$EventType <- as.character(resistance_total$EventType)
resistance_total$Location <- as.character(resistance_total$Location)
resistance_total$EventYear <- as.character(resistance_total$EventYear)
resistance_total$AgeCategory <- as.character(resistance_total$AgeCategory)

# recode state to abbrev
resistance_total <- resistance_total %>%
  mutate(State = recode(Location, 
                           "Alabama" = "AL",
                           "Alaska" = "AK",
                           "Arizona" = "AZ",
                           "Arkansas" = "AR",
                           "California" = "CA",
                           "Colorado" = "CO",
                           "Connecticut" = "CT",
                           "Delaware" = "DE",
                           "District Of Columbia" = "DC",
                           "Florida" = "FL",
                           "Georgia" = "GA",
                           "Hawaii" = "HI",
                           "Idaho" = "ID",
                           "Illinois" = "IL",
                           "Indiana" = "IN", 
                           "Iowa" = "IA",
                           "Kansas" = "KS",
                           "Kentucky" = "KY",
                           "Louisiana" = "LA", 
                           "Maine" = "ME",
                           "Maryland" = "MD",
                           "Massachusetts" = "MA",
                           "Michigan" = "MI",
                           "Minnesota" = "MN",
                           "Mississippi" = "MS",
                           "Missouri" = "MO",
                           "Montana" = "MT",
                           "Nebraska" = "NE",
                           "Nevada" = "NV",
                           "New Hampshire" = "NH",
                           "New Jersey" = "NJ",
                           "New Mexico" = "NM",
                           "New York" = "NY",
                           "North Carolina" = "NC",
                           "North Dakota" = "ND",
                           "Ohio" = "OH",
                           "Oklahoma" = "OK",
                           "Oregon" = "OR",
                           "Pennsylvania" = "PA",
                           "Rhode Island" = "RI",
                           "South Carolina" = "SC",
                           "South Dakota" = "SD",
                           "Tennessee" = "TN",
                           "Texas" = "TX",
                           "Utah" = "UT",
                           "Vermont" = "VT",
                           "Virginia" = "VA",
                           "Washington" = "WA",
                           "West Virginia" = "WV",
                           "Wisconsin" = "WI",
                           "Wyoming" = "WY")) %>%
  mutate(Region = ifelse(State %in% c("IL", "IN", "MI", "OH", "WI"), "East North Central",
                         ifelse(State %in% c("AL", "KY", "MS", "TN"), "East South Central",
                                ifelse(State %in% c("NJ", "NY", "PA"), "Middle Atlantic",
                                       ifelse(State %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY"), "Mountain",
                                              ifelse(State %in% c("CT", "ME", "MA", "NH", "RI", "VT"), "New England",
                                                     ifelse(State %in% c("AK", "CA", "HI", "OR", "WA"), "Pacific",
                                                            ifelse(State %in% c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV"), "South Atlantic",
                                                                   ifelse(State %in% c("IA", "KS", "MN", "MO", "NE", "ND", "SD"), "West North Central",
                                                                          ifelse(State %in% c("AR", "LA", "OK", "TX"), "West South Central", "Region Undefined"))))))))))

# reordering
col_order <- c("Phenotype", "EventType", "Location", "State", "Region", "EventYear", "AgeCategory", "numberTested", "numberResistant", "pop_count", "agedistribution")

resistance_total <- resistance_total[, col_order]

# Finding the means of numberTested and number Resistant by REGION and AGE:

summary_region_mean <- resistance_total %>%
  group_by(Region, AgeCategory) %>%
  summarize(AvgNumTested = mean(numberTested, na.rm = TRUE),
            AvgNumResistant = mean(numberResistant, na.rm = TRUE))



# With plyr package:
plyr_resistance_total <- ddply(resistance_total,~ Phenotype  +  EventType + Location + EventYear + AgeCategory,
      summarise,average=mean(numberTested)) 

```












