---
title: "Special_Studies"
output: html_document
---

# Data Loading:
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)

# 1) RESISTANCE DATA
resistance_state <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_States.csv")

# 2) 
resistance_region <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_Regional.csv")

# 3)
resistance_national <- read.csv("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Resist_National.csv")

# 4) OUPATIENT ANTIBIOTIC USE DATA
antibiotic_use <- read.csv ("~/Documents/School/Smith Sem 7/Special Studies/PSP_Antibiotic_Resistance_Data/Outpatient_Antibiotic_Use.csv") %>%
  mutate(State = recode(Location, 
                           "Alabama" = "AL",
                           "Alaska" = "AK",
                           "Arizona" = "AZ",
                           "Arkansas" = "AR",
                           "California" = "CA",
                           "Colorado" = "CO",
                           "Connecticut" = "CT",
                           "Delaware" = "DE",
                           "District Of Columbia" = "DC",
                           "Florida" = "FL",
                           "Georgia" = "GA",
                           "Hawaii" = "HI",
                           "Idaho" = "ID",
                           "Illinois" = "IL",
                           "Indiana" = "IN", 
                           "Iowa" = "IA",
                           "Kansas" = "KS",
                           "Kentucky" = "KY",
                           "Louisiana" = "LA", 
                           "Maine" = "ME",
                           "Maryland" = "MD",
                           "Massachusetts" = "MA",
                           "Michigan" = "MI",
                           "Minnesota" = "MN",
                           "Mississippi" = "MS",
                           "Missouri" = "MO",
                           "Montana" = "MT",
                           "Nebraska" = "NE",
                           "Nevada" = "NV",
                           "New Hampshire" = "NH",
                           "New Jersey" = "NJ",
                           "New Mexico" = "NM",
                           "New York" = "NY",
                           "North Carolina" = "NC",
                           "North Dakota" = "ND",
                           "Ohio" = "OH",
                           "Oklahoma" = "OK",
                           "Oregon" = "OR",
                           "Pennsylvania" = "PA",
                           "Rhode Island" = "RI",
                           "South Carolina" = "SC",
                           "South Dakota" = "SD",
                           "Tennessee" = "TN",
                           "Texas" = "TX",
                           "Utah" = "UT",
                           "Vermont" = "VT",
                           "Virginia" = "VA",
                           "Washington" = "WA",
                           "West Virginia" = "WV",
                           "Wisconsin" = "WI",
                           "Wyoming" = "WY"))

```


## New Data:
```{r}
## Aggregate Data:
collapse_AntibioticUse <- antibiotic_use %>%
  filter(Location != "National" & Antibiotic_Class == "All classes") %>%
  mutate(State = recode(Location, 
                           "Alabama" = "AL",
                           "Alaska" = "AK",
                           "Arizona" = "AZ",
                           "Arkansas" = "AR",
                           "California" = "CA",
                           "Colorado" = "CO",
                           "Connecticut" = "CT",
                           "Delaware" = "DE",
                           "District Of Columbia" = "DC",
                           "Florida" = "FL",
                           "Georgia" = "GA",
                           "Hawaii" = "HI",
                           "Idaho" = "ID",
                           "Illinois" = "IL",
                           "Indiana" = "IN", 
                           "Iowa" = "IA",
                           "Kansas" = "KS",
                           "Kentucky" = "KY",
                           "Louisiana" = "LA", 
                           "Maine" = "ME",
                           "Maryland" = "MD",
                           "Massachusetts" = "MA",
                           "Michigan" = "MI",
                           "Minnesota" = "MN",
                           "Mississippi" = "MS",
                           "Missouri" = "MO",
                           "Montana" = "MT",
                           "Nebraska" = "NE",
                           "Nevada" = "NV",
                           "New Hampshire" = "NH",
                           "New Jersey" = "NJ",
                           "New Mexico" = "NM",
                           "New York" = "NY",
                           "North Carolina" = "NC",
                           "North Dakota" = "ND",
                           "Ohio" = "OH",
                           "Oklahoma" = "OK",
                           "Oregon" = "OR",
                           "Pennsylvania" = "PA",
                           "Rhode Island" = "RI",
                           "South Carolina" = "SC",
                           "South Dakota" = "SD",
                           "Tennessee" = "TN",
                           "Texas" = "TX",
                           "Utah" = "UT",
                           "Vermont" = "VT",
                           "Virginia" = "VA",
                           "Washington" = "WA",
                           "West Virginia" = "WV",
                           "Wisconsin" = "WI",
                           "Wyoming" = "WY")) %>%
  select(Year, Rate, State)

collapse_resistance <- resistance_state %>%
  filter(NumberResistant != "Insufficient Data" & NumberResistant != "None Tested") %>%
  filter(EventYear != "All Years") %>%
  filter(AgeCategory == "All Ages") %>%
  rename(Year = "EventYear") %>%
  mutate(Region = ifelse(State %in% c("IL", "IN", "MI", "OH", "WI"), "East North Central",
                         ifelse(State %in% c("AL", "KY", "MS", "TN"), "East South Central",
                                ifelse(State %in% c("NJ", "NY", "PA"), "Middle Atlantic",
                                       ifelse(State %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY"), "Mountain",
                                              ifelse(State %in% c("CT", "ME", "MA", "NH", "RI", "VT"), "New England",
                                                     ifelse(State %in% c("AK", "CA", "HI", "OR", "WA"), "Pacific",
                                                            ifelse(State %in% c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV"), "South Atlantic",
                                                                   ifelse(State %in% c("IA", "KS", "MN", "MO", "NE", "ND", "SD"), "West North Central",
                                                                          ifelse(State %in% c("AR", "LA", "OK", "TX"), "West South Central", "Region Undefined"))))))))))

collapse_resistance$NumberTested <- as.numeric(as.character(collapse_resistance$NumberTested))
collapse_resistance$NumberResistant <- as.numeric(as.character(collapse_resistance$NumberResistant))
collapse_resistance$PercentResistant <- as.numeric(as.character(collapse_resistance$PercentResistant))
collapse_AntibioticUse$Rate <- as.numeric(as.character(collapse_AntibioticUse$Rate))

collapse_resistance <- collapse_resistance %>%
  select(State, Region, Year,  NumberTested, NumberResistant, PercentResistant) %>%
  group_by(State, Region, Year) %>%
  summarise(NumberTested = sum(NumberTested),
            NumberResistant = sum(NumberResistant),
            PercentResistant = sum(PercentResistant)) %>%
  mutate(calculated_percent = 100*(NumberResistant/NumberTested))

```

```{r}
# Separate by year:

## 2011:
Resistance_2011 <- collapse_resistance %>%
  filter(Year == "2011")
Antibiotic_2011 <- collapse_AntibioticUse %>%
  filter(Year == "2011")
joined_2011 <- Antibiotic_2011 %>%
  inner_join(Resistance_2011, by = "State")

## 2012:
Resistance_2012 <- collapse_resistance %>%
  filter(Year == "2012")
Antibiotic_2012 <- collapse_AntibioticUse %>%
  filter(Year == "2012")
joined_2012 <- Antibiotic_2012 %>%
  inner_join(Resistance_2012, by = "State")

## 2013:
Resistance_2013 <- collapse_resistance %>%
  filter(Year == "2013")
Antibiotic_2013 <- collapse_AntibioticUse %>%
  filter(Year == "2013")
joined_2013 <- Antibiotic_2013 %>%
  inner_join(Resistance_2013, by = "State")

## 2014:
Resistance_2014 <- collapse_resistance %>%
  filter(Year == "2014")
Antibiotic_2014 <- collapse_AntibioticUse %>%
  filter(Year == "2014")
joined_2014 <- Antibiotic_2014 %>%
  inner_join(Resistance_2014, by = "State")
```

## Plotting year:
```{r}
# 2011:
## Using Percent Resistant
#ggplot(data = joined_2011, mapping = aes(x = Rate, y = PercentResistant, color = Region)) + geom_point() + labs(title = "2011 Percent Resistant vs. Rate of Antibiotic Use")
## Using Calculated Percent
ggplot(data = joined_2011, mapping = aes(x = Rate, y = calculated_percent, color = Region)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use") + scale_color_manual(values =c('red','orange', 'yellow', 'green', 'blue', 'purple', 'grey', 'brown', 'black'))

# 2012:
## Using Percent Resistant
#ggplot(data = joined_2012, mapping = aes(x = Rate, y = PercentResistant, color = Region)) + geom_point() + labs(title = "2012 Percent Resistant vs. Rate of Antibiotic Use")
## Using Calculated Percent
ggplot(data = joined_2012, mapping = aes(x = Rate, y = calculated_percent, color = Region)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use") + scale_color_manual(values =c('red','orange', 'yellow', 'green', 'blue', 'purple', 'grey', 'brown', 'black'))

# 2013:
## Using Percent Resistant
#ggplot(data = joined_2013, mapping = aes(x = Rate, y = PercentResistant, color = Region)) + geom_point() + labs(title = "2013 Percent Resistant vs. Rate of Antibiotic Use")
## Using Calculated Percent
ggplot(data = joined_2013, mapping = aes(x = Rate, y = calculated_percent, color = Region)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use") + scale_color_manual(values =c('red','orange', 'yellow', 'green', 'blue', 'purple', 'grey', 'brown', 'black'))

# 2014:
## Using Percent Resistant
#ggplot(data = joined_2014, mapping = aes(x = Rate, y = PercentResistant, color = Region)) + geom_point() + labs(title = "2014 Percent Resistant vs. Rate of Antibiotic Use")
## Using Calculated Percent
ggplot(data = joined_2014, mapping = aes(x = Rate, y = calculated_percent, color = Region)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use") + scale_color_manual(values =c('red','orange', 'yellow', 'green', 'blue', 'purple', 'grey', 'brown', 'black'))

regmod <- lm(calculated_percent ~ Rate, data = joined_2011)
summary(regmod)
regmod <- lm(calculated_percent ~ Rate, data = joined_2012)
summary(regmod)
regmod <- lm(calculated_percent ~ Rate, data = joined_2013)
summary(regmod)
regmod <- lm(calculated_percent ~ Rate, data = joined_2014)
summary(regmod)
```


## Regional Parsing 
```{r}

# EastNorthCentral - 2011
EastNorthCentral_2011 <- joined_2011 %>%
  filter(Region == "East North Central")

# EastSouthCentral - 2011
EastSouthCentral_2011 <- joined_2011 %>%
  filter(Region == "East South Central")

# Middle Atlantic - 2011
MiddleAtlantic_2011 <- joined_2011 %>%
  filter(Region == "Middle Atlantic")

# Mountain - 2011
Mountain_2011 <- joined_2011 %>%
  filter(Region == "Mountain")

# New England - 2011
NewEngland_2011 <- joined_2011 %>%
  filter(Region == "New England")

# Pacific - 2011
Pacific_2011 <- joined_2011 %>%
  filter(Region == "Pacific")

# South Atlantic - 2011
SouthAtlantic_2011 <- joined_2011 %>%
  filter(Region == "South Atlantic")

# West North Central - 2011
WestNorthCentral_2011 <- joined_2011 %>%
  filter(Region == "West North Central")

# West South Central - 2011
WestSouthCentral_2011 <- joined_2011 %>%
  filter(Region == "West South Central")

#########################################

# EastNorthCentral - 2012
EastNorthCentral_2012 <- joined_2012 %>%
  filter(Region == "East North Central")

# EastSouthCentral - 2012
EastSouthCentral_2012 <- joined_2012 %>%
  filter(Region == "East South Central")

# Middle Atlantic - 2012
MiddleAtlantic_2012 <- joined_2012 %>%
  filter(Region == "Middle Atlantic")

# Mountain - 2012
Mountain_2012 <- joined_2012 %>%
  filter(Region == "Mountain")

# New England - 2012
NewEngland_2012 <- joined_2012 %>%
  filter(Region == "New England")

# Pacific - 2012
Pacific_2012 <- joined_2012 %>%
  filter(Region == "Pacific")

# South Atlantic - 2012
SouthAtlantic_2012 <- joined_2012 %>%
  filter(Region == "South Atlantic")

# West North Central - 2012
WestNorthCentral_2012 <- joined_2012 %>%
  filter(Region == "West North Central")

# West South Central - 2012
WestSouthCentral_2012 <- joined_2012 %>%
  filter(Region == "West South Central")

##########################################

# EastNorthCentral - 2013
EastNorthCentral_2013 <- joined_2013 %>%
  filter(Region == "East North Central")

# EastSouthCentral - 2013
EastSouthCentral_2013 <- joined_2013 %>%
  filter(Region == "East South Central")

# Middle Atlantic - 2013
MiddleAtlantic_2013 <- joined_2013 %>%
  filter(Region == "Middle Atlantic")

# Mountain - 2013
Mountain_2013 <- joined_2013 %>%
  filter(Region == "Mountain")

# New England - 2013
NewEngland_2013 <- joined_2013 %>%
  filter(Region == "New England")

# Pacific - 2013
Pacific_2013 <- joined_2013 %>%
  filter(Region == "Pacific")

# South Atlantic - 2013
SouthAtlantic_2013 <- joined_2013 %>%
  filter(Region == "South Atlantic")

# West North Central - 2013
WestNorthCentral_2013 <- joined_2013 %>%
  filter(Region == "West North Central")

# West South Central - 2013
WestSouthCentral_2013 <- joined_2013 %>%
  filter(Region == "West South Central")


#########################################

# EastNorthCentral - 2014
EastNorthCentral_2014 <- joined_2014 %>%
  filter(Region == "East North Central")

# EastSouthCentral - 2014
EastSouthCentral_2014 <- joined_2014 %>%
  filter(Region == "East South Central")

# Middle Atlantic - 2014
MiddleAtlantic_2014 <- joined_2014 %>%
  filter(Region == "Middle Atlantic")

# Mountain - 2014
Mountain_2014 <- joined_2014 %>%
  filter(Region == "Mountain")

# New England - 2014
NewEngland_2014 <- joined_2014 %>%
  filter(Region == "New England")

# Pacific - 2014
Pacific_2014 <- joined_2014 %>%
  filter(Region == "Pacific")

# South Atlantic - 2014
SouthAtlantic_2014 <- joined_2014 %>%
  filter(Region == "South Atlantic")

# West North Central - 2014
WestNorthCentral_2014 <- joined_2014 %>%
  filter(Region == "West North Central")

# West South Central - 2014
WestSouthCentral_2014 <- joined_2014 %>%
  filter(Region == "West South Central")



```

## Regional Graph:
```{r}
ggplot(data = EastNorthCentral_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for East North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "red")

ggplot(data = EastSouthCentral_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for East South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "orange")

ggplot(data = MiddleAtlantic_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for Middle Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "yellow")

ggplot(data = Mountain_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for Mountain Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "green")

ggplot(data = NewEngland_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for New England Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "blue")

ggplot(data = Pacific_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for Pacific Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "purple")

ggplot(data = SouthAtlantic_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2011 Calculated Resistant vs. Rate of Antibiotic Use for South Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "grey")

ggplot(data = WestNorthCentral_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for West North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "brown")

ggplot(data = WestSouthCentral_2011, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for West South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#############################################

ggplot(data = EastNorthCentral_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for East North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "red")

ggplot(data = EastSouthCentral_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for East South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "orange")

ggplot(data = MiddleAtlantic_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for Middle Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "yellow")

ggplot(data = Mountain_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for Mountain Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "green")

ggplot(data = NewEngland_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for New England Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "blue")

ggplot(data = Pacific_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for Pacific Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "purple")

ggplot(data = SouthAtlantic_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for South Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "grey")

ggplot(data = WestNorthCentral_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for West North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "brown")

ggplot(data = WestSouthCentral_2012, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2012 Calculated Resistant vs. Rate of Antibiotic Use for West South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##########################################

ggplot(data = EastNorthCentral_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for East North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "red")

ggplot(data = EastSouthCentral_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for East South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "orange")

ggplot(data = MiddleAtlantic_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for Middle Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "yellow")

ggplot(data = Mountain_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for Mountain Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "green")

ggplot(data = NewEngland_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for New England Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "blue")

ggplot(data = Pacific_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for Pacific Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "purple")

ggplot(data = SouthAtlantic_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for South Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "grey")

ggplot(data = WestNorthCentral_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for West North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "brown")

ggplot(data = WestSouthCentral_2013, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2013 Calculated Resistant vs. Rate of Antibiotic Use for West South Central Region")  + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#######################################################

ggplot(data = EastNorthCentral_2014, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for East North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "red")

ggplot(data = EastSouthCentral_2014, mapping = aes(x = Rate, y = calculated_percent)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for East South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "orange")

ggplot(data = MiddleAtlantic_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for Middle Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "yellow")

ggplot(data = Mountain_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for Mountain Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "green")

ggplot(data = NewEngland_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for New England Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "blue")

ggplot(data = Pacific_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for Pacific Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "purple")

ggplot(data = SouthAtlantic_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for South Atlantic Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "grey")

ggplot(data = WestNorthCentral_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for West North Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "brown")

ggplot(data = WestSouthCentral_2014, mapping = aes(x = Rate, y = calculated_percent, size = NumberTested)) + geom_point() + labs(title = "2014 Calculated Resistant vs. Rate of Antibiotic Use for West South Central Region") + xlim(450, 1400) + ylim(0,30) + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")


```



## Only MDR for Acinetobacter spp.: 
```{r}
# resistance state:
OnlyMDR <- resistance_state %>%
  filter(EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs")

# Acinetobacter spp. MDR and Cephalosporins
ceph_use <- antibiotic_use %>%
  select(Antibiotic_Class, Year, Rate, State) %>%
  filter(Antibiotic_Class != "All classes", Antibiotic_Class != "All other", 
         State != "National", Antibiotic_Class == "Cephalosporins", Year < "2015") %>%
  group_by(Antibiotic_Class, State) %>%
  summarise(Total_Rate = sum(as.numeric(as.character(Rate)), na.rm = TRUE))

MDR_Acinetobacter <- OnlyMDR %>%
  filter(Phenotype == "Acinetobacter spp. MDR")

MDR_Acin_ceph <- ceph_use %>%
  inner_join(MDR_Acinetobacter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_ceph, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. MDR and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Acin_ceph$PercentResistant <- as.numeric(as.character(MDR_Acin_ceph$PercentResistant))
MDR_Acin_ceph$Total_Rate <- as.numeric(as.character(MDR_Acin_ceph$Total_Rate))

r = cor(MDR_Acin_ceph$Total_Rate, MDR_Acin_ceph$PercentResistant)
r

##############

# Acinetobacter spp. MDR and Fluoroquinolones

fluo_use <- antibiotic_use %>%
  select(Antibiotic_Class, Year, Rate, State) %>%
  filter(Antibiotic_Class != "All classes", Antibiotic_Class != "All other", 
         State != "National", Antibiotic_Class == "Fluoroquinolones", Year < "2015") %>%
  group_by(Antibiotic_Class, State) %>%
  summarise(Total_Rate = sum(as.numeric(as.character(Rate)), na.rm = TRUE))

MDR_Acin_fluo <- fluo_use %>%
  inner_join(MDR_Acinetobacter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant, NumberTested) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_fluo, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant), size = as.numeric(as.character(NumberTested))))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. MDR and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Acin_fluo$PercentResistant <- as.numeric(as.character(MDR_Acin_fluo$PercentResistant))
MDR_Acin_fluo$Total_Rate <- as.numeric(as.character(MDR_Acin_fluo$Total_Rate))

r = cor(MDR_Acin_fluo$Total_Rate, MDR_Acin_fluo$PercentResistant)
r

##############

# Acinetobacter spp. MDR and Macrolides

macro_use <- antibiotic_use %>%
  select(Antibiotic_Class, Year, Rate, State) %>%
  filter(Antibiotic_Class != "All classes", Antibiotic_Class != "All other", 
         State != "National", Antibiotic_Class == "Macrolides", Year < "2015") %>%
  group_by(Antibiotic_Class, State) %>%
  summarise(Total_Rate = sum(as.numeric(as.character(Rate)), na.rm = TRUE))

MDR_Acin_macro <- macro_use %>%
  inner_join(MDR_Acinetobacter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_macro, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. MDR and Macrolides Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Acin_macro$PercentResistant <- as.numeric(as.character(MDR_Acin_macro$PercentResistant))
MDR_Acin_macro$Total_Rate <- as.numeric(as.character(MDR_Acin_macro$Total_Rate))

r = cor(MDR_Acin_macro$Total_Rate, MDR_Acin_macro$PercentResistant)
r

##############

# Acinetobacter spp. MDR and Penicillins

pen_use <- antibiotic_use %>%
  select(Antibiotic_Class, Year, Rate, State) %>%
  filter(Antibiotic_Class != "All classes", Antibiotic_Class != "All other", 
         State != "National", Antibiotic_Class == "Penicillins", Year < "2015") %>%
  group_by(Antibiotic_Class, State) %>%
  summarise(Total_Rate = sum(as.numeric(as.character(Rate)), na.rm = TRUE))

MDR_Acin_pen <- pen_use %>%
  inner_join(MDR_Acinetobacter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_pen, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. MDR and Penicillins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Acin_pen$PercentResistant <- as.numeric(as.character(MDR_Acin_pen$PercentResistant))
MDR_Acin_pen$Total_Rate <- as.numeric(as.character(MDR_Acin_pen$Total_Rate))

r = cor(MDR_Acin_pen$Total_Rate, MDR_Acin_pen$PercentResistant)
r

```


## Only MDR for E.coli:
```{r}
# E.coli MDR and Cephalosporins
MDR_Ecoli <- OnlyMDR %>%
  filter(Phenotype == "E.coli MDR")

MDR_Ecoli_ceph <- ceph_use %>%
  inner_join(MDR_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_ceph, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 E.coli MDR and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Ecoli_ceph$PercentResistant <- as.numeric(as.character(MDR_Ecoli_ceph$PercentResistant))
MDR_Ecoli_ceph$Total_Rate <- as.numeric(as.character(MDR_Ecoli_ceph$Total_Rate))

r = cor(MDR_Ecoli_ceph$Total_Rate, MDR_Ecoli_ceph$PercentResistant)
r

##############

# E.coli MDR and Fluoroquinolones

MDR_Ecoli_fluo <- fluo_use %>%
  inner_join(MDR_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Ecoli_fluo, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 E.coli MDR and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Ecoli_fluo$PercentResistant <- as.numeric(as.character(MDR_Ecoli_fluo$PercentResistant))
MDR_Ecoli_fluo$Total_Rate <- as.numeric(as.character(MDR_Ecoli_fluo$Total_Rate))

r = cor(MDR_Ecoli_fluo$Total_Rate, MDR_Ecoli_fluo$PercentResistant)
r #0.349

##############

# E.coli MDR and Macrolides

MDR_Ecoli_macro <- macro_use %>%
  inner_join(MDR_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Ecoli_macro, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 E.coli MDR and Macrolides Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Ecoli_macro$PercentResistant <- as.numeric(as.character(MDR_Ecoli_macro$PercentResistant))
MDR_Ecoli_macro$Total_Rate <- as.numeric(as.character(MDR_Ecoli_macro$Total_Rate))

r = cor(MDR_Ecoli_macro$Total_Rate, MDR_Ecoli_macro$PercentResistant)
r

##############

# E.coli MDR and Penicillins

MDR_Ecoli_pen <- pen_use %>%
  inner_join(MDR_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Ecoli_pen, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 E.coli MDR and Penicillins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

# r coeff
MDR_Ecoli_pen$PercentResistant <- as.numeric(as.character(MDR_Ecoli_pen$PercentResistant))
MDR_Ecoli_pen$Total_Rate <- as.numeric(as.character(MDR_Ecoli_pen$Total_Rate))

r = cor(MDR_Ecoli_pen$Total_Rate, MDR_Ecoli_pen$PercentResistant)
r

```


## Only MDR for Enterobacter spp.:
```{r}
# Enterobacter spp. MDR and Cephalosporins
MDR_Enter <- OnlyMDR %>%
  filter(Phenotype == "Enterobacter spp. MDR")

MDR_Enter_ceph <- ceph_use %>%
  inner_join(MDR_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Enter_ceph, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. MDR and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Enterobacter spp. MDR and Fluoroquinolones

MDR_Enter_fluo <- fluo_use %>%
  inner_join(MDR_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Enter_fluo, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. MDR and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Enterobacter spp. MDR and Macrolides

MDR_Enter_macro <- macro_use %>%
  inner_join(MDR_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Enter_macro, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. MDR and Macrolides Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Enterobacter spp. MDR and Penicillins

MDR_Enter_pen <- pen_use %>%
  inner_join(MDR_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Enter_pen, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. MDR and Penicillins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```


## Only MDR for Klebsiella spp.:
```{r}
# Klebsiella spp. MDR and Cephalosporins
MDR_Kleb <- OnlyMDR %>%
  filter(Phenotype == "Klebsiella spp. MDR")

MDR_Kleb_ceph <- ceph_use %>%
  inner_join(MDR_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Kleb_ceph, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. MDR and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Klebsiella spp. MDR and Fluoroquinolones

MDR_Kleb_fluo <- fluo_use %>%
  inner_join(MDR_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Kleb_fluo, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. MDR and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Klebsiella spp. MDR and Macrolides

MDR_Kleb_macro <- macro_use %>%
  inner_join(MDR_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Kleb_macro, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. MDR and Macrolides Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Klebsiella spp. MDR and Penicillins

MDR_Kleb_pen <- pen_use %>%
  inner_join(MDR_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Kleb_pen, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. MDR and Penicillins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```


## Only MDR for P. aeruginosa:
```{r}
# P. aeruginosa MDR and Cephalosporins
MDR_Paeru <- OnlyMDR %>%
  filter(Phenotype == "P. aeruginosa MDR")

MDR_Paeru_ceph <- ceph_use %>%
  inner_join(MDR_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Paeru_ceph, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa MDR and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#############

# P. aeruginosa MDR and Fluoroquinolones

MDR_Paeru_fluo <- fluo_use %>%
  inner_join(MDR_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Paeru_fluo, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa MDR and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# P. aeruginosa MDR and Macrolides

MDR_Paeru_macro <- macro_use %>%
  inner_join(MDR_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Paeru_macro, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa MDR and Macrolides Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# P. aeruginosa MDR and Penicillins

MDR_Paeru_pen <- pen_use %>%
  inner_join(MDR_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Paeru_pen, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa MDR and Penicillins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```

## MDR Bacteria with All Classes:
```{r}
All_use <- antibiotic_use %>%
  select(Antibiotic_Class, Year, Rate, State) %>%
  filter(Antibiotic_Class == "All classes", State != "National", Year < "2015") %>%
  group_by(Antibiotic_Class, State) %>%
  summarise(Total_Rate = sum(as.numeric(as.character(Rate)), na.rm = TRUE))

##############

# Acinetobacter spp. MDR and All Classes

MDR_Acin_All <- All_use %>%
  inner_join(MDR_Acinetobacter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Acin_All, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. MDR and All Class Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# E.coli MDR and All Classses

MDR_Ecoli_All <- All_use %>%
  inner_join(MDR_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Ecoli_All, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 E.coli MDR and All Class Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

###############

# Enterobacter spp. MDR and All Classses

MDR_Enter_All <- All_use %>%
  inner_join(MDR_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Enter_All, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. MDR and All Class Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# Klebsiella spp. MDR and All Classses

MDR_Kleb_All <- All_use %>%
  inner_join(MDR_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Kleb_All, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. MDR and All Class Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

##############

# P. aeruginosa MDR and All Classses

MDR_Paeru_All <- All_use %>%
  inner_join(MDR_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, Phenotype, PercentResistant) %>%
  filter(PercentResistant != "Insufficient Data", PercentResistant != "Not Defined")

ggplot(data = MDR_Paeru_All, mapping = aes(x = Total_Rate, y = as.numeric(as.character(PercentResistant)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa MDR and All Class Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")



```

# 	Bacteria MDR + R:
## 	P. aeruginosa:
```{r}
#	P. aeruginosa MDR and R to ES cephalosporins AND Cephalosporins:
Both_MDRCeph_Paeru <- resistance_state %>%
  filter(Phenotype == "P. aeruginosa MDR" | Phenotype == "P. aeruginosa R to ES cephalosporins", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRCeph_Paeru <- ceph_use %>%
  inner_join(Both_MDRCeph_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRCeph_Paeru, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa (MDR +R) and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	P. aeruginosa MDR and R to Fluoroquinolones AND Fluoroquinolones:
Both_MDRfluo_Paeru <- resistance_state %>%
  filter(Phenotype == "P. aeruginosa MDR" | Phenotype == "P. aeruginosa R to fluroquinolones", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRfluo_Paeru <- fluo_use %>%
  inner_join(Both_MDRfluo_Paeru, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRfluo_Paeru, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 P. aeruginosa (MDR + R) and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```

## E.coli MDR + R:
```{r}
#	Ecoli MDR and R to ES cephalosporins AND Cephalosporins:
Both_MDRCeph_Ecoli <- resistance_state %>%
  filter(Phenotype == "E.coli MDR" | Phenotype == "E.coli R to ES cephalosporins", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRCeph_Ecoli <- ceph_use %>%
  inner_join(Both_MDRCeph_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRCeph_Ecoli, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 E.coli (MDR +R) and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	Ecoli MDR and R to Fluoroquinolones AND Fluoroquinolones:
Both_MDRFluo_Ecoli <- resistance_state %>%
  filter(Phenotype == "E.coli MDR" | Phenotype == "E.coli R to fluoroquinolones	", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRFluo_Ecoli <- fluo_use %>%
  inner_join(Both_MDRFluo_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRFluo_Ecoli, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 E.coli (MDR +R) and Fluoroquinolones Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```


## Enterobacter spp. MDR + R:
```{r}
#	Enterobacter spp. MDR and R to ES cephalosporins AND Cephalosporins:
Both_MDRCeph_Enter <- resistance_state %>%
  filter(Phenotype == "Enterobacter spp. MDR" | Phenotype == "Enterobacter spp. R to ES cephalosporins", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRCeph_Enter <- ceph_use %>%
  inner_join(Both_MDRCeph_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRCeph_Enter, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. (MDR +R) and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```


## Klebsiella spp. MDR + R:
```{r}
#	Klebsiella spp. MDR and R to ES cephalosporins AND Cephalosporins:
Both_MDRCeph_Kleb <- resistance_state %>%
  filter(Phenotype == "Klebsiella spp. MDR" | Phenotype == "Klebsiella spp. R to ES cephalosporins", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

MDRCeph_Kleb <- ceph_use %>%
  inner_join(Both_MDRCeph_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = MDRCeph_Kleb, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. (MDR +R) and Cephalosporins Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")
```


# Excluding MDR for all Bacteria:
```{r}
# All Acinetobacter spp. Resistance (excluding MDR)

#	Acinetobacter spp. (R + I) AND All Use:
All_Acin <- resistance_state %>%
  filter(Phenotype == "Acinetobacter spp. (R+I) to carbapenems", 
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

All_R_Acin <- All_use %>%
  inner_join(All_Acin, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = All_R_Acin, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Acinetobacter spp. (R + I) and All Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	E.coli R AND All Use:
All_Ecoli <- resistance_state %>%
  filter(Phenotype == "E.coli R to carbapenems" |  Phenotype == "E.coli R to ES cephalosporins" | 
           Phenotype == "E.coli R to fluoroquinolones",
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

All_R_Ecoli <- All_use %>%
  inner_join(All_Ecoli, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = All_R_Ecoli, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 E.coli R and All Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	Enterobacter spp. R AND All Use:
All_Enter <- resistance_state %>%
  filter(Phenotype == "Enterobacter spp. R to carbapenems" |  Phenotype == "Enterobacter spp. R to ES cephalosporins",
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

All_R_Enter <- All_use %>%
  inner_join(All_Enter, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = All_R_Enter, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Enterobacter spp. R and All Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	Klebsiella spp. R AND All Use:
All_Kleb <- resistance_state %>%
  filter(Phenotype == "Klebsiella spp. R to carbapenems" |  Phenotype == "Klebsiella spp. R to ES cephalosporins",
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

All_R_Kleb <- All_use %>%
  inner_join(All_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = All_R_Kleb, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. R and All Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

#	Klebsiella spp. R AND All Use:
All_Kleb <- resistance_state %>%
  filter(Phenotype == "Klebsiella spp. R to carbapenems" |  Phenotype == "Klebsiella spp. R to ES cephalosporins",
         EventYear == "All Years", AgeCategory == "All Ages", EventType == "All HAIs") %>%
  select(-Phenotype) %>%
  mutate(CalculatedResistance = ((as.numeric(as.character(NumberResistant))/as.numeric(as.character(NumberTested)))*100)) %>%
  group_by(State) %>%
  summarise(SumCalculatedResistance = sum(CalculatedResistance))

All_R_Kleb <- All_use %>%
  inner_join(All_Kleb, by = "State") %>%
  select(Antibiotic_Class, State, Total_Rate, SumCalculatedResistance)

ggplot(data = All_R_Kleb, mapping = aes(x = Total_Rate, y = as.numeric(as.character(SumCalculatedResistance)))) + geom_point() + labs(title = "2011-2014 Klebsiella spp. R and All Use") + geom_smooth(method = "lm", se = FALSE, fullrange = TRUE, color = "black")

```

# Specific antibiotic use class to resistance
```{r}
# Cephalasporin:

```




